Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2016-11-13 manuscript.tex
latest

On the other hand, the model does remarkably well in capturing apparently general features of selection under different environmental conditions using only three trait parameters: $b$, $c$ and $d$. The clean separation of a strictly-relative $c$ parameter is particularly useful from an evolutionary genetics perspective, essentially embedding a relative fitness trait within an absolute fitness model. This could have interesting applications for modeling the impacts of intra-specific competition on species extinction, without arbitrarily assigning an absolute fitness cost to that competition. [citations to sexual selection stuff?]

\begin{figure}
\centering
\begin{tabularx}{\linewidth}{lXXXXX}
& Emeddable in flexible genetic scheme? & Ecologically meaningful? & Empirically-grounded trait scheme? & Generality beyond specific scenarios? & Formal model? \\ \hline
Density-dependent lottery & \ding{51} & \ding{51} & \ding{51} & \ding{51} & \ding{51} \\  
  MacArthur's $r$/$K$ + $\alpha$ & \ding{51} & \ding{51} & \ding{55}* & \ding{51} & \ding{51} \\
  Grime's C/S/R & NA & \ding{51} & \ding{51} & \ding{51} & \ding{55} \\
  Traditional pop. gen. & \ding{51} & \ding{55} & \ding{55} & \ding{51} & \ding{51} \\
  Eco-evo:  \\
  Adaptive dynamics & \ding{55} & \ding{51} & \ding{51} & \ding{55}** & \ding{51} \\
  Brute force simulation & \ding{51} & \ding{51} & NA & \ding{55} & \ding{51}
\end{tabularx}
\caption{\label{fig:table2} Comparison of our density-dependent lottery model with related models/schemes in ecology and evolutionary biology.}
\end{figure}

>>>>

On the other hand, our model does remarkably well in capturing apparently general features of selection under different environmental conditions using only three trait parameters: $b$, $c$ and $d$. The clean separation of a strictly-relative $c$ parameter is particularly useful from an evolutionary genetics perspective, essentially embedding a relative fitness trait within an absolute fitness model. This could have interesting applications for modeling the impacts of intra-specific competition on species extinction, without arbitrarily assigning an absolute fitness cost to that competition. [citations to sexual selection stuff?]

In the broader context of eco-evolutionary synthesis [? Cite], we can imagine a spectrum of approaches ranging from population genetics, which is entirely evolutionary (no ecology) and has a strong emphasis on genetics, to pure ecological theory with no evolution.

\begin{figure}
\centering
\begin{tabularx}{\linewidth}{lXXXXX}
&Formal model?  & Ecologically meaningful? & Empirically-grounded trait scheme? & Generality beyond specific scenarios? & Genetically flexible? \\ \hline
Density-dependent lottery & \ding{51} & \ding{51} & \ding{51} & \ding{51} & \ding{51} \\  
  MacArthur's $r$/$K$ + $\alpha$ & \ding{51} & \ding{51} & \ding{55}* & \ding{51} & \ding{51} \\
  Grime's C/S/R & \ding{55} & \ding{51} & \ding{51} & \ding{51} & NA \\
  Traditional pop. gen. & \ding{51} & \ding{55} & \ding{55} & \ding{51} & \ding{51} \\
  Eco-evo:  \\
  \quad Adaptive dynamics & \ding{51} & \ding{51} & \ding{51} & \ding{55}** & \ding{55} \\
  \quad Brute force simulation & \ding{51} & \ding{51} & NA & \ding{55} & \ding{51}
\end{tabularx}
\caption{\label{fig:table2} Comparison of our density-dependent lottery model with related models and schemes in ecology and evolutionary biology. *MacArthur's $r$- and $K$-, as well as $\alpha$ selection, were all derived theoretically. Applications to traits come later, and with mixed success (see ``Introduction'' and ``Discussion''). **As a modeling framework, adaptive dynamics is not scenario-specific, and could be applied with any fitness model including our density-dependent lottery, but in practice most of the adaptive dynamics literature focuses on specific eco-evolutionary outcomes such as evolutionary ``branching'' or evolutionary suicide.}
\end{figure}
